---
title: "FLACC results"
---
Load libraries
```{r}
library(ggplot2)
library(dplyr)
library(FSA)
library(ggbeeswarm)
library(stringr)
library(pheatmap)
library(Matrix)
```
```{r colour bar, echo=FALSE, fig.height=2, fig.width=13}
palette <- c(
 # Writers
  "MTA"     = "#004488",
  "VIR"  = "#009988",
  "FIP37"   = "#0077BB",
  "hiz2-4" = "#AA4499",
  "hiz2-2"    = "#CCAA22",
  "HAKAI" = "#556B2F",
  #Readers
  "ECT"     = "#EE7733",
  "CPSF30"  = "#DD3377",
  #Other
  "U2B"     = "#CC3311",
  "Col-0"    = "#888888"
)
barplot(rep(1, 10), col = palette, names.arg = names(palette), yaxt = "n")
```
Load genome and annotation
```{r}
At.genome <- Rsamtools::FaFile("/Volumes/cluster/ggs_lab/cmetheringham001/Shared_Ref/TAIR10_chr_all.fas")
# Load TxDb
txdb <- makeTxDbFromGFF("/Volumes/cluster/ggs_lab/cmetheringham001/Shared_Ref/Araport11_GFF3_genes_transposons.201606.no_chr.gtf")
```
Plot the proportion of A in DRACH motifs
```{r}
drach_summary <- data.frame(geno = c(rep("Col0", 4), rep("flacc_mutant", 4), rep("flacc_ox", 4)),
                            prop = c(0.0507, 0.0481, 0.0482, 0.0462,
                                     0.0090, 0.0060, 0.0066, 0.0069,
                                     0.0531, 0.0508, 0.0524, 0.0529))
```
Plot in same style as LC-MS/MS plot
```{r}
custom_colors <- c("Col0" = "black", "flacc_mutant" = "#CCAA22", "flacc_ox" = "#AA4499")
custom_shapes <- c("Col0" = 16, "flacc_mutant" = 15, "flacc_ox" = 18)

# Create the plot
summary_plot <- ggplot(data = drach_summary, aes(y = prop, x = geno, colour = geno, shape = geno)) +
  geom_boxplot(colour = "darkgrey", outlier.shape = NA) +
  geom_beeswarm(size = 3, cex = 2) +
  theme_classic(base_size = 12) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    limits = c(0, NA)
  ) +
  scale_x_discrete(
    labels = c(
      "Col0" = "Col-0",
      "flacc_mutant" = expression(italic("hiz2-2")),
      "flacc_ox" = expression(italic("hiz2-4"))
    )
  ) +
  labs(
    x = "Genotype",
    y = "Proportion on DRACH predicted modified",
    colour = "Genotype",
    shape = "Genotype"
  ) +
  scale_color_manual(
    values = custom_colors,
    labels = c(
      "Col0" = "Col-0",
      "flacc_mutant" = expression(italic("hiz2-2")),
      "flacc_ox" = expression(italic("hiz2-4"))
    )
  ) +
  scale_shape_manual(
    values = custom_shapes,
    labels = c(
      "Col0" = "Col-0",
      "flacc_mutant" = expression(italic("hiz2-2")),
      "flacc_ox" = expression(italic("hiz2-4"))
    )
  ) +
  guides(
    colour = guide_legend(title = "Genotype"),
    shape = guide_legend(title = "Genotype")
  )

summary_plot
```
```{r}
ggsave(
  filename = "../../Figures/figure_2/rna004_flacc_m6a_prediction.pdf",
  plot = summary_plot,
  width = 5, height = 4, units = "in"
)
ggsave(
  filename = "../../Figures/figure_2/rna004_flacc_m6a_prediction.png",
  plot = summary_plot,
  width = 5, height = 4, units = "in" 
)
```

Is this a significant difference?
```{r}
drach_summary_filtered <- drach_summary %>% filter(!is.na(prop))
shapiro.test(drach_summary_filtered$prop)
kruskal.test(prop ~ geno, data = drach_summary_filtered)
dunnTest(prop ~ geno, data = drach_summary_filtered, method = "bonferroni")
t.test(c(0.0507, 0.0481, 0.0482, 0.0462), c(0.0090, 0.0060, 0.0066, 0.0069), conf.int = 0.95)
t.test(c(0.0531, 0.0508, 0.0524, 0.0529), c(0.0507, 0.0481, 0.0482, 0.0462), conf.int = 0.95)
```
Modifications on a per site basis
```{r}
flacc_mutant_1 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/flacc_mutant_1.bed", header=FALSE)
flacc_mutant_2 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/flacc_mutant_2.bed", header=FALSE)
flacc_mutant_3 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/flacc_mutant_3.bed", header=FALSE)
flacc_mutant_4 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/flacc_mutant_4.bed", header=FALSE)
```

```{r}
flacc_OX_1 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/flacc_OX_1.bed", header=FALSE)
flacc_OX_2 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/flacc_OX_2.bed", header=FALSE)
flacc_OX_3 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/flacc_OX_3.bed", header=FALSE)
flacc_OX_4 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/flacc_OX_4.bed", header=FALSE)
```
```{r}
col0_1 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/col0_1.bed", header=FALSE)
col0_2 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/col0_2.bed", header=FALSE)
col0_3 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/col0_3.bed", header=FALSE)
col0_4 <- read.delim("~/Documents/Reader_Writer_Paper/SIKFLACC/modkit/col0_4.bed", header=FALSE)
```
Get colnames
```{r}
bed_colnames <- c(
  "chrom",
  "start",
  "end",
  "mod_code_motif",
  "score",
  "strand",
  "start_compat",
  "end_compat",
  "color",
  "Nvalid_cov",
  "percent_modified",
  "Nmod",
  "Ncanonical",
  "Nother_mod",
  "Ndelete",
  "Nfail",
  "Ndiff",
  "Nnocall"
)
set_colnames <- function(df_list, new_colnames) {
  lapply(df_list, function(df) {
    colnames(df) <- new_colnames
    return(df)
  })
}
```
Make list and rename
```{r}
bed_list <- list(
  col0_1 = col0_1,
  col0_2 = col0_2,
  col0_3 = col0_3,
  col0_4 = col0_4,
  flacc_mutant_1 = flacc_mutant_1,
  flacc_mutant_2 = flacc_mutant_2,
  flacc_mutant_3 = flacc_mutant_3,
  flacc_mutant_4 = flacc_mutant_4,
  flacc_OX_1 = flacc_OX_1,
  flacc_OX_2 = flacc_OX_2,
  flacc_OX_3 = flacc_OX_3,
  flacc_OX_4 = flacc_OX_4
)
bed_list <- set_colnames(bed_list, bed_colnames)
```

Filter the sites
```{r}
filter_modification_tables <- function(df_list, min_percent = 0.01, min_coverage = 20) {
  filtered_list <- lapply(names(df_list), function(name) {
    df <- df_list[[name]]
    df_filtered <- df[df$percent_modified > min_percent & df$Nvalid_cov >= min_coverage, ]
    return(df_filtered)
  })
  names(filtered_list) <- paste0(names(df_list), ".filtered")
  return(filtered_list)
}
filtered_bed_tables <- filter_modification_tables(bed_list)
```

Where are these sites located?
```{r}
# Extract UTRs
five_utrs <- fiveUTRsByTranscript(txdb, use.names=TRUE)
three_utrs <- threeUTRsByTranscript(txdb, use.names=TRUE)
# Flatten into GRanges
five_utrs_flat <- unlist(five_utrs, use.names=FALSE)
three_utrs_flat <- unlist(three_utrs, use.names=FALSE)
five_utrs_flat$region_type <- "5UTR"
three_utrs_flat$region_type <- "3UTR"
# Combine 5' and 3' UTRs into one GRanges
utr_regions <- c(five_utrs_flat, three_utrs_flat)
```
Annotate with UTR
```{r}
annotate_with_UTRs <- function(df_list, utr_regions) {
  annotated_list <- lapply(names(df_list), function(name) {
    df <- df_list[[name]]
    # Create GRanges object from the input table
    data_gr <- GRanges(
      seqnames = df$chrom,
      ranges = IRanges(start = df$start, end = df$end),
      strand = df$strand,
      transcript_id = df$name
    )
    # Find overlaps with UTR regions
    hits <- findOverlaps(data_gr, utr_regions, ignore.strand = FALSE)
    # Annotate with UTR region type
    utr_annotation <- rep(NA_character_, length(data_gr))
    utr_annotation[queryHits(hits)] <- mcols(utr_regions)$region_type[subjectHits(hits)]
    # Add UTR_type column to data frame
    annotated_df <- df %>%
      mutate(UTR_type = utr_annotation)
    return(annotated_df)
  })
  # Name output list with ".annotated" suffix
  names(annotated_list) <- paste0(names(df_list), ".annotated")
  return(annotated_list)
}
annotated_filtered_bed_tables <- annotate_with_UTRs(filtered_bed_tables, utr_regions)
```
Summarise the UTR types
```{r}
summarise_utr_percentages <- function(df_list) {
  utr_summary <- lapply(names(df_list), function(name) {
    df <- df_list[[name]]
    # Check
    if (!"UTR_type" %in% colnames(df)) {
      return(data.frame(File = name, UTR_type = NA, Percent = NA))
    }
    # Tabulate and calculate percentages
    total <- nrow(df)
    tab <- prop.table(table(df$UTR_type)) * 100  
    df_tab <- as.data.frame(tab)
    colnames(df_tab) <- c("UTR_type", "Percent")
    df_tab$File <- name
    return(df_tab)
  })
  # Combine into data frame
  do.call(rbind, utr_summary) %>%
    dplyr::select(File, UTR_type, Percent)
}
utr_percent_table <- summarise_utr_percentages(annotated_filtered_bed_tables)
range(utr_percent_table$Percent[utr_percent_table$UTR_type=="3UTR"])
range(utr_percent_table$Percent[utr_percent_table$UTR_type=="5UTR"])
```

How do these sites overlap with 002 results?
```{r}
hiz2_4_diff_err <- fread("/Volumes/cluster/ggs_lab/mtparker/papers/flacc/flacc_nanopore/pipeline/differr/flacc_vs_col0.bed")
colnames(hiz2_4_diff_err) <- c("chrom", "start", "end", "name", "score", "strand", "odds", "G", "P", "FDR", "g_hom", "P_hom", "g_hom_c", "P_hom_c")
vir1_diff_err <- fread("/Volumes/cluster/ggs_lab/mtparker/papers/flacc/flacc_nanopore/pipeline/differr/vir1_vs_VIRc.bed")
colnames(vir1_diff_err) <- c("chrom", "start", "end", "name", "score", "strand", "odds", "G", "P", "FDR", "g_hom", "P_hom", "g_hom_c", "P_hom_c")
head(hiz2_4_diff_err)
```
Calculate overlaps
```{r}
calculate_overlap_proportion <- function(df_list, gr1_source, maxgap = 5, ignore_strand = FALSE) {
  # Create gr1 from the fixed input
  gr1 <- GRanges(
    seqnames = gr1_source$chrom,
    ranges = IRanges(start = gr1_source$start, end = gr1_source$end),
    strand = gr1_source$strand
  )
  total <- length(gr1)
  proportions <- sapply(names(df_list), function(name) {
    df <- df_list[[name]]
    gr2 <- GRanges(
      seqnames = df$chrom,
      ranges = IRanges(start = df$start, end = df$end),
      strand = df$strand
    )
    hits <- findOverlaps(gr1, gr2, maxgap = maxgap, ignore.strand = ignore_strand)
    n_hits <- length(unique(queryHits(hits)))
    return(n_hits / total)
  })
  return(proportions)
}
```

Number of flacc diff_err sites which were modified in the new data
```{r}
hiz2_overlap_props <- calculate_overlap_proportion(filtered_bed_tables, hiz2_4_diff_err)
hiz2_overlap_props
vir1_overlap_props <- calculate_overlap_proportion(filtered_bed_tables, vir1_diff_err)
vir1_overlap_props
```
Make a summary table
```{r}
summarise_table_list <- function(df_list) {
  key_cols <- c("chrom", "start", "end", "mod_code_motif", "strand")
  
  processed_list <- lapply(names(df_list), function(name) {
    df <- df_list[[name]]
    
    # Trim the table name
    short_name <- str_remove(name, "\\.filtered.*")  # removes ".filtered.annotated" etc
    
    df %>%
      dplyr::select(all_of(c(key_cols, "Nvalid_cov", "percent_modified"))) %>%
      rename(
        !!paste0(short_name, ".Nvalid_cov") := Nvalid_cov,
        !!paste0(short_name, ".percent_modified") := percent_modified
      )
  })
  
  # Merge all tables by the key columns
  merged_df <- Reduce(function(x, y) full_join(x, y, by = key_cols), processed_list)
  
  return(merged_df)
}
summary_table <- summarise_table_list(bed_list)
head(summary_table)
```
Compute the per site means
```{r}
summary_table_means <- summary_table %>%
  mutate(
    mean_Nvalid_cov = rowMeans(dplyr::select(., ends_with(".Nvalid_cov")), na.rm = TRUE),
    mean_percent_modified = rowMeans(dplyr::select(., ends_with(".percent_modified")), na.rm = TRUE),

    mean_Nvalid_cov_col0 = rowMeans(dplyr::select(., matches("^col0.*\\.Nvalid_cov$")), na.rm = TRUE),
    mean_percent_modified_col0 = rowMeans(dplyr::select(., matches("^col0.*\\.percent_modified$")), na.rm = TRUE),

    mean_Nvalid_cov_mutant = rowMeans(dplyr::select(., matches("^flacc_mutant.*\\.Nvalid_cov$")), na.rm = TRUE),
    mean_percent_modified_mutant = rowMeans(dplyr::select(., matches("^flacc_mutant.*\\.percent_modified$")), na.rm = TRUE),

    mean_Nvalid_cov_OX = rowMeans(dplyr::select(., matches("^flacc_OX.*\\.Nvalid_cov$")), na.rm = TRUE),
    mean_percent_modified_OX = rowMeans(dplyr::select(., matches("^flacc_OX.*\\.percent_modified$")), na.rm = TRUE)
  )
head(summary_table_means)
```
Filter the table by the mean coverage
```{r}
summary_table_means.filtered <- summary_table_means[summary_table_means$mean_Nvalid_cov>=20,]
summary_table_means.filtered <- summary_table_means.filtered[summary_table_means$mean_percent_modified_col0>0,]
# write to table
write_csv(summary_table_means.filtered, file = "flacc004_modifications.csv")
```
Plot the comparison of the mean percentage modified Col-0 and hiz2-2
```{r}
t.test(summary_table_means.filtered$mean_percent_modified_col0, summary_table_means.filtered$mean_percent_modified_mutant, conf.level = 0.95)
t.test(summary_table_means.filtered$mean_percent_modified_col0, summary_table_means.filtered$mean_percent_modified_OX, conf.level = 0.95)
```

What sites are there that are unique to each?
```{r}
summary_table.mods <- summary_table[, grep("\\.percent_modified$", names(summary_table))]
#summary_table.mods <- summary_table.mods[summary_table_means$mean_Nvalid_cov>=20,]
summary_table.mods <- summary_table.mods[(summary_table_means$mean_Nvalid_cov_col0>=20 &
                                              summary_table_means$mean_Nvalid_cov_mutant>=20 &
                                              summary_table_means$mean_Nvalid_cov_OX>=20),]
#summary_table.mods <- summary_table.mods[rowSums(summary_table.mods != 0, na.rm = TRUE) > 0, ]
summary_table.mods <- summary_table.mods[rowSums(summary_table.mods > 8, na.rm = TRUE) > 0, ]
colnames(summary_table.mods) <- sub("\\.percent_modified$", "", colnames(summary_table.mods))
head(summary_table.mods)
```
Find columns that are unique to flacc_OX in this table
```{r}
# Identify col0 and flacc_OX columns
col0_cols <- grep("^col0_", names(summary_table.mods), value = TRUE)
flacc_OX_cols <- grep("^flacc_OX_", names(summary_table.mods), value = TRUE)

threshold <- 1
unique_flacc_OX <- summary_table.mods[
  rowSums(summary_table.mods[col0_cols] > threshold) == 0 &
  rowSums(summary_table.mods[flacc_OX_cols] > threshold) > 0,
]
unique_col0 <- summary_table.mods[
  rowSums(summary_table.mods[flacc_OX_cols] > threshold) == 0 &
  rowSums(summary_table.mods[col0_cols] > threshold) > 0,
]
```

Venn diagram of the overlap
```{r}
library(eulerr)
# Define the sets
fit <- euler(c(
  "Col-0" = 192 ,   # unique + overlap
  "hiz2-4" = 447 ,
  "Col-0&hiz2-4" = 21087
))
# Plot with custom colours
plot(fit,
     fills = list(fill = c("gray30", "#AA4499"), alpha = 0.6),
     edges = "gray30",
     labels = list(col = "black", font = 2),
     quantities = TRUE)
```
Calculate the means of the modified data
```{r}
summary_table.mods.means <- summary_table.mods %>%
  rowwise() %>%
  mutate(
    mean_col0 = mean(c_across(starts_with("col0_")), na.rm = TRUE),
    mean_mutant = mean(c_across(starts_with("flacc_mutant_")), na.rm = TRUE),
    mean_OX = mean(c_across(starts_with("flacc_OX_")), na.rm = TRUE)
  )
head(summary_table.mods.means)
```

Plot the comparison of the mean percentage modified Col-0 and hiz2-2
```{r}
df2 <- summary_table.mods.means
df2$higher_in_x <- df2$mean_col0 < df2$mean_mutant
df2$lower_in_x <- df2$mean_col0 > df2$mean_mutant

hiz2_2_sites <- ggplot(df2, aes(x = mean_col0,
               y = mean_mutant,
               color = lower_in_x)) +
  geom_point(alpha = 0.3, show.legend = TRUE) +  
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_manual(values = c("TRUE" = "#CCAA22", "FALSE" = "grey50")) +
  theme_minimal(base_size = 10) +
  labs(
    x = "Mean % Modified (Col-0)",
    y = "Mean % Modified (hiz2-2)",
    size = "Mean Coverage"
  ) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    legend.bmutant = "horizontal"
  ) +
  guides(
    color = "none",          
    size = "none"
  )

hiz2_2_sites
ggsave(
  filename = "../../Figures/figure_2/hiz2_2_sites.pdf",
  plot = hiz2_2_sites,
  width = 5, height = 4, units = "in"
)
ggsave(
  filename = "../../Figures/figure_2/hiz2_2_sites.png",
  plot = hiz2_2_sites,
  width = 5, height = 4, units = "in" 
)
table(df2$higher_in_x)
table(df2$lower_in_x)
```
Plot the comparison of the mean percentage modified Col-0 and hiz2-4
```{r}
df <- summary_table.mods.means
df$higher_in_x <- df$mean_col0 < df$mean_OX
df$lower_in_x <- df$mean_col0 > df$mean_OX

hiz2_4_sites <- ggplot(df, aes(x = mean_col0,
               y = mean_OX,
               color = higher_in_x)) +
  geom_point(alpha = 0.3, show.legend = TRUE) +  
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  scale_color_manual(values = c("TRUE" = "#AA4499", "FALSE" = "grey50")) +
  theme_minimal(base_size = 10) +
  labs(
    x = "Mean % Modified (Col-0)",
    y = "Mean % Modified (hiz2-4)",
    size = "Mean Coverage"
  ) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    legend.box = "horizontal"
  ) +
  guides(
    color = "none",          
    size = "none"
 )

hiz2_4_sites
ggsave(
  filename = "../../Figures/figure_2/hiz2_4_sites.pdf",
  plot = hiz2_4_sites,
  width = 5, height = 4, units = "in"
)
ggsave(
  filename = "../../Figures/figure_2/hiz2_4_sites.png",
  plot = hiz2_4_sites,
  width = 5, height = 4, units = "in" 
)
table(df$higher_in_x)
table(df$lower_in_x)
```
Fit a linear model
```{r}
library(car)
model <- lm(mean_OX ~ mean_col0, data = summary_table.mods.means)
summary(model)
linearHypothesis(model, "mean_col0 = 1") 
linearHypothesis(model, c("(Intercept) = 0", "mean_col0 = 1"))
linearHypothesis(model, c("(Intercept) = 0"))
```
Plot a heatmap
```{r}
summary_table.mods[] <- lapply(summary_table.mods, as.numeric)
# Create sparse matrix
mat_sparse <- as(Matrix(as.matrix(summary_table.mods), sparse = TRUE), "dgCMatrix")
# Convert the entire sparse matrix to dense (no sampling)
mat_dense <- as.matrix(mat_sparse)
# Replace NA/NaN/Inf with 0
mat_dense[!is.finite(mat_dense)] <- 0
# Optional: set rownames (if you have meaningful row IDs in summary_table.mods)
rownames(mat_dense) <- rownames(summary_table.mods)
```
Plot the heatmap
```{r}
# Plot heatmap with pheatmap
heatmap <- pheatmap(mat_dense,
         scale = "none",
         show_rownames = FALSE,
         labels_col = c("Col-0 1", "Col-0 2", "Col-0 3", "Col-0 4",
                        expression(italic("hiz2-4")*1), expression(italic("hiz2-4")*2), expression(italic("hiz2-4")*3), expression(italic("hiz2-4")*4),
                         expression(italic("hiz2-2")*1), expression(italic("hiz2-2")*2), expression(italic("hiz2-2")*3), expression(italic("hiz2-2")*4)),
         color = colorRampPalette(c("white", "firebrick4"))(50))
heatmap
```
Save the heatmap to file
```{r}
ggsave(
  filename = "../../Figures/figure_2/m6a_heatmap.pdf",
  plot = heatmap,
  width = 5, height = 5, units = "in"
)
ggsave(
  filename = "../../Figures/figure_2/m6a_heatmap.png",
  plot = heatmap,
  width = 5, height = 5, units = "in" 
)
```

