---
title: "Proteomics Plots"
---
```{r}
library(data.table)
library(readr)
library(ggplot2)
library(cowplot)
library(discretization)
library(ggrepel)
library(stringr)
library(eulerr)
library(org.At.tair.db)
library(AnnotationDbi)
library(readxl)
library(dplyr)
```
Colour palette
```{r colour bar, echo=FALSE, fig.height=2, fig.width=13}
palette <- c(
 # Writers
  "MTA"     = "#004488",
  "VIR"  = "#009988",
  "FIP37"   = "#0077BB",
  "hiz2-4" = "#AA4499",
  "hiz2-2"    = "#CCAA22",
  "HAKAI" = "#556B2F",

  #Readers
  "ECT"     = "#EE7733",
  "CPSF30"  = "#DD3377",
  
  #Other
  "U2B"     = "#CC3311",
  "?"    = "#BBBBBB"
)
barplot(rep(1, 10), col = palette, names.arg = names(palette), yaxt = "n")
```


Read in the proteomics data
```{r}
ect_vs_col0 <- fread("/Volumes/cluster/ggs_lab/cmetheringham001/Readers_Writers/proteomics/proteus_output/ect2_vs_col0.tsv")
hakai_vs_col0 <- fread("/Volumes/cluster/ggs_lab/cmetheringham001/Readers_Writers/proteomics/proteus_output/hakai_vs_col0.tsv")
u2b_vs_col0 <- fread("/Volumes/cluster/ggs_lab/cmetheringham001/Readers_Writers/proteomics/proteus_output/u2b_vs_col0.tsv")
vir_vs_col0 <- fread("/Volumes/cluster/ggs_lab/cmetheringham001/Readers_Writers/proteomics/proteus_output/vir1_vs_col0.tsv")
mta_vs_col0 <- fread("/Volumes/cluster/ggs_lab/cmetheringham001/Readers_Writers/proteomics/proteus_output/mta_vs_col0.tsv")
mta_vs_mtahakai <- fread("/Volumes/cluster/ggs_lab/cmetheringham001/Readers_Writers/proteomics/proteus_output/MTA_vs_MTAhakai.tsv")
```

Tidy proteomics data
```{r}
tables <- list(
  ect_vs_col0 = ect_vs_col0,
  hakai_vs_col0 = hakai_vs_col0,
  mta_vs_mtahakai = mta_vs_mtahakai,
  mta_vs_col0 = mta_vs_col0,
  vir_vs_col0 = vir_vs_col0,
  u2b_vs_col0 = u2b_vs_col0
)
for (name in names(tables)) {
    tables[[name]][, protein := sapply(strsplit(protein, ".", fixed = TRUE), `[`, 1)]
}
```

Get all gene ID and GO terms
```{r}
for (name in names(tables)) {
    gene_names <- AnnotationDbi::select(org.At.tair.db,
                                    keys = tables[[name]]$protein,
                                    columns = c("SYMBOL"),
                                    keytype = "TAIR")
    table <- merge(tables[[name]], gene_names, by.x = "protein", by.y = "TAIR", all.x = TRUE)
    table_merged <- table[, .(
    SYMBOL = paste(unique(na.omit(SYMBOL)), collapse = ",")
  ), by = eval(names(table)[1:11])]
    #Save
    tables[[name]] <- table_merged
    write_csv(table_merged, file = paste(name, "csv", sep = "."))
}
```
Read in the spreadsheet with Gordon's annotations
```{r}
excel_file <- "../../Proteomics/GS edit U2B_MTA_VIR_FPA_proteomics_results.xlsx"
sheet_names <- excel_sheets(excel_file)
# Read and extract the first 3 columns from each sheet
sheet_data_list <- lapply(sheet_names, function(sheet) {
  read_excel(excel_file, sheet = sheet) %>%
    select(1:5)
})

# Optionally, name the list elements by sheet name
names(sheet_data_list) <- sheet_names
u2b_info <- sheet_data_list[["U2B_vs_Col-0"]]
```


Protein lists
```{r}
mrna_interacting <- fread("../../Proteomics/mrna_interacting.txt")
ects <- fread("../../Proteomics/ect_proteins.txt")
writer_complex <- c("AT5G01160", "AT3G05680", "AT5G53440", "AT4G10760", "AT4G09980", "AT3G54170")
```
Annotations from Gordon's list

LogFC plots

#U2BL
```{r}
u2b_vs_col02 <- merge(u2b_vs_col0, u2b_info, by.x = "protein", by.y = "agi_id", all.x = T)

u2b_vs_col02$colour <- u2b_vs_col02$category
u2b_vs_col02$colour[u2b_vs_col02$colour == "other"] <- NA
# Label U2B
u2b_vs_col02$colour[u2b_vs_col02$protein == "AT1G06960"] <- 'U2B\"'
u2b_vs_col02$name[u2b_vs_col02$protein == "AT1G06960"] <- 'U2B\"'

# Define color vector with matching names
category_colors <- c(
  'U2B"' = "#CC3311",
  "U2_snRNP" = "#CC79A7",
  "splicing_factor" = "#009E73",
  "other_rna_binding" = "#0072B2"
)
new_labels <- c(
  'U2B"' = 'U2B"',
  "U2_snRNP" = "U2 snRNP",
  "splicing_factor" = "Splicing factor",
  "other_rna_binding" = "Other RNA binding",
  "NA" = "Other"
)

desired_order <- c('U2B"', "U2_snRNP", "splicing_factor", "other_rna_binding")
u2b_vs_col02$colour <- factor(u2b_vs_col02$colour, levels = desired_order)

plot_u2b_vs_col0 <- ggplot(data = u2b_vs_col02[u2b_vs_col02$U2B>=3,], aes(y=-log10(pval), x=median_logFC, label = name, fill = colour, size = median_av_exprs))+
    geom_point(shape = 21,colour = "white")+
    theme_classic(base_size=15)+
    geom_text_repel(aes(label=name, hjust=1.2), size = 6, max.overlaps = Inf)+
    xlab("median logFC")+
    #xlim(-4,12)+
    xlab('Median log2 Fold Change')+
    ylab('-log10 p value')+
    ggtitle('U2B" IP')+
    scale_fill_manual(
        values = category_colors,
        labels = new_labels, 
        name = "Category"
    ) +
    theme(legend.position = c(.2, .7))+
    guides(size = "none", fill = guide_legend(title = "Category"))

plot_u2b_vs_col0

ggsave(
  filename = "../../Figures/figure_1/u2b_vs_col0_ip.pdf",
  plot = plot_u2b_vs_col0,
  width = 6, height = 5, units = "in" 
)
ggsave(
  filename = "../../Figures/figure_1/u2b_vs_col0_ip.png",
  plot = plot_u2b_vs_col0,
  width = 6, height = 5, units = "in" 
)
```

#HAKAI
```{r}
writer_palette <- c(
  "MTA"     = "#004488",
  "MTB"     = "#004488",
  "VIR"     = "#009988",
  "FIP37"   = "#0077BB",
  "HIZ2"  = "#CCAA22",
  "HAKAI"   = "#556B2F",
  "HIZ1" = "#CC3311",
  "mRNA interacting" = "purple"
)

hakai_vs_col0$name[hakai_vs_col0$protein == "AT5G01160"] <- "HAKAI"
hakai_vs_col0$name[hakai_vs_col0$protein == "AT3G05680"] <- "VIR"
hakai_vs_col0$name[hakai_vs_col0$protein == "AT5G53440"] <- "HIZ2"
hakai_vs_col0$name[hakai_vs_col0$protein == "AT4G10760"] <- "MTA"
hakai_vs_col0$name[hakai_vs_col0$protein == "AT4G09980"] <- "MTB"
hakai_vs_col0$name[hakai_vs_col0$protein == "AT3G54170"] <- "FIP37"
hakai_vs_col0$name[hakai_vs_col0$protein == "AT1G32360"] <- "HIZ1"
hakai_vs_col0$colour <- writer_palette[hakai_vs_col0$name]
hakai_vs_col0$colour[is.na(hakai_vs_col0$colour)] <- "#B5B5B5"

plot_hakai_vs_col0 <- ggplot(data = hakai_vs_col0[hakai_vs_col0$HAKAI >= 3,], 
                             aes(y = -log10(pval), x = median_logFC, label = name, fill = colour, size = median_av_exprs)) +
  geom_point(shape = 21, colour = "white") +
  theme_classic(base_size = 15) +
  geom_text_repel(aes(label = name, hjust = 1.2), size = 4, max.overlaps = Inf) +
  xlab('Median log2 Fold Change') +
  ylab('-Log10 p value') +
  ggtitle("HAKAI IP") +
  scale_fill_identity(guide = "none") +
  theme(legend.position = "none") +
  guides(size = "none", colour = "none")

plot_hakai_vs_col0

ggsave(
  filename = "../../Figures/figure_1/hakai_vs_col0_ip.pdf",
  plot = plot_hakai_vs_col0,
  width = 6, height = 5, units = "in" 
)
ggsave(
  filename = "../../Figures/figure_1/hakai_vs_col0_ip.png",
  plot = plot_hakai_vs_col0,
  width = 6, height = 5, units = "in" 
)
```
#MTA
```{r}
mta_vs_col0$name[mta_vs_col0$protein == "AT5G01160"] <- "HAKAI"
mta_vs_col0$name[mta_vs_col0$protein == "AT3G05680"] <- "VIR"
mta_vs_col0$name[mta_vs_col0$protein == "AT5G53440"] <- "HIZ2"
mta_vs_col0$name[mta_vs_col0$protein == "AT4G10760"] <- "MTA"
mta_vs_col0$name[mta_vs_col0$protein == "AT4G09980"] <- "MTB"
mta_vs_col0$name[mta_vs_col0$protein == "AT3G54170"] <- "FIP37"
mta_vs_col0$name[mta_vs_col0$protein == "AT1G32360"] <- "HIZ1"

mta_vs_col0$colour <- writer_palette[mta_vs_col0$name]
mta_vs_col0$colour[is.na(mta_vs_col0$colour)] <- "#B5B5B5"

plot_mta_vs_col0 <- ggplot(data = mta_vs_col0[mta_vs_col0$MTA>=3,], aes(y=-log10(pval), x=median_logFC, label = name, fill = colour, size = median_av_exprs))+
  geom_point(shape = 21, colour = "white") +
  theme_classic(base_size = 15) +
  geom_text_repel(aes(label = name, hjust = 1.3), size = 4, max.overlaps = Inf) +
  xlab('Median log2 Fold Change') +
  ylab('-Log10 p value') +
  ggtitle("MTA vs Col0 IP") +
  scale_fill_identity(guide = "none") +
  theme(legend.position = "none") +
  guides(size = "none", colour = "none")

plot_mta_vs_col0

ggsave(
  filename = "../../Figures/figure_1/mta_vs_col0_ip.pdf",
  plot = plot_mta_vs_col0,
  width = 6, height = 5, units = "in" 
)
ggsave(
  filename = "../../Figures/figure_1/mta_vs_col0_ip.png",
  plot = plot_mta_vs_col0,
  width = 6, height = 5, units = "in" 
)
```
#VIR

```{r}
vir1_vs_col0$name <- NA_character_
vir1_vs_col0$name[vir1_vs_col0$protein %in% mrna_interacting$protein] <- "mRNA interacting"
vir1_vs_col0$name[vir1_vs_col0$protein == "AT5G01160"] <- "HAKAI"
vir1_vs_col0$name[vir1_vs_col0$protein == "AT3G05680"] <- "VIR"
vir1_vs_col0$name[vir1_vs_col0$protein == "AT5G53440"] <- "HIZ2"
vir1_vs_col0$name[vir1_vs_col0$protein == "AT4G10760"] <- "MTA"
vir1_vs_col0$name[vir1_vs_col0$protein == "AT4G09980"] <- "MTB"
vir1_vs_col0$name[vir1_vs_col0$protein == "AT3G54170"] <- "FIP37"
vir1_vs_col0$name[vir1_vs_col0$protein == "AT1G32360"] <- "HIZ1"

# Define palette with purple for mRNA interacting
writer_palette_extended <- c(
  writer_palette,
  "mRNA interacting" = "#7f47ca"  
)

# Assign colours based on name, default grey if missing
vir1_vs_col0$colour <- writer_palette_extended[vir1_vs_col0$name]
vir1_vs_col0$colour[is.na(vir1_vs_col0$colour)] <- "#B5B5B5"

# Create a logical to label only non-mRNA interacting proteins
vir1_vs_col0$label_plot <- ifelse(vir1_vs_col0$name == "mRNA interacting", NA, vir1_vs_col0$name)

# Plot
plot_vir1_vs_col0 <- ggplot(data = vir1_vs_col0[vir1_vs_col0$VIR >= 3,], 
                            aes(y = -log10(pval), x = median_logFC, label = label_plot, fill = colour, size = median_av_exprs)) +
  geom_point(shape = 21, colour = "white") +
  theme_classic(base_size = 15) +
  geom_text_repel(aes(label = label_plot, hjust = 1.3), size = 4, max.overlaps = Inf, na.rm = TRUE) +
  xlab('Median log2 Fold Change') +
  ylab('-Log10 p value') +
  ggtitle("VIR vs Col0 IP") +
  scale_fill_identity(
    guide = "legend",
    breaks = "#7f47ca",
    labels = "mRNA interacting",
    name = ""
  ) +
  theme(legend.position = c(0.8, 0.2)) +
  guides(size = "none", colour = "none")+
  # Add purple dot as key 
  annotate("point", x = max(vir1_vs_col0$median_logFC, na.rm = TRUE) * 0.6, 
                     y = max(-log10(vir1_vs_col0$pval), na.rm = TRUE) * 0.1,
                     colour = "#7f47ca", size = 3, shape = 21, fill = "#7f47ca") +
  # Add text label next to the dot
  annotate("text", x = max(vir1_vs_col0$median_logFC, na.rm = TRUE) * 0.65,
                   y = max(-log10(vir1_vs_col0$pval), na.rm = TRUE) * 0.1,
                   label = "mRNA interacting", hjust = 0, size = 4)

plot_vir1_vs_col0

ggsave(
  filename = "../../Figures/figure_1/vir_vs_col0_ip.pdf",
  plot = plot_vir1_vs_col0,
  width = 6, height = 5, units = "in" 
)
ggsave(
  filename = "../../Figures/figure_1/vir_vs_col0_ip.png",
  plot = plot_vir1_vs_col0,
  width = 6, height = 5, units = "in" 
)
```

Overlaps between the 3 sets
```{r}
mta <- mta_vs_col0[pval < 0.05 & median_logFC > 1 & MTA>=3, protein]
vir1 <- vir1_vs_col0[pval < 0.05 & median_logFC > 1 & VIR>=3, protein]
hakai <- hakai_vs_col0[pval < 0.05 & median_logFC > 1 & HAKAI>=3, protein]
u2b <- u2b_vs_col0[pval < 0.05 & median_logFC > 1 & U2B>=3, protein]

venn_list <- euler(list(
  MTA = mta,
  VIR = vir1,
  HAKAI = hakai
))

venn_plot <- plot(venn_list,
     fills = list(fill = c("#004488", "#009988", "#556B2F"), alpha = 0.7),
     labels = list(font = 2, cex = 1.2),
     edges = TRUE,
     quantities = TRUE)

venn_plot

ggsave(
  filename = "../../Figures/figure_1/mta_vir_hakai_overlap.pdf",
  plot = venn_plot,
  width = 6, height = 5, units = "in" 
)
ggsave(
  filename = "../../Figures/figure_1/mta_vir_hakai_overlap.png",
  plot = venn_plot,
  width = 6, height = 5, units = "in" 
)
```
The 5 proteins in the overlap are HIZ1, VIR, FIP37, MTA and HAKAI

```{r}
all_proteins <- unique(c(mta, vir1, hakai))

overlap_dt <- data.table(
  protein = all_proteins,
  MTA = all_proteins %in% mta,
  VIR = all_proteins %in% vir1,
  HAKAI = all_proteins %in% hakai
)

# View table
overlap_dt

# Optional: subset proteins in all three
overlap_dt[MTA & VIR & HAKAI]

# Get mappings
gene_names <- AnnotationDbi::select(org.At.tair.db,
                                    keys = overlap_dt$protein,
                                    columns = c("SYMBOL"),
                                    keytype = "TAIR")

# Merge
overlap_dt <- merge(overlap_dt, gene_names, by.x = "protein", by.y = "TAIR", all.x = TRUE)
overlap_dt <- overlap_dt[, .(
  SYMBOL = paste(unique(SYMBOL), collapse = ",")
), by = .(protein, MTA, VIR, HAKAI)]

head(overlap_dt)
write_csv(overlap_dt, file = "protein_overlap.csv")
```
Overlap between VIR and U2B
```{r}
venn_list <- euler(list(
  U2B = u2b,
  VIR = vir1
))

plot(venn_list,
     fills = list(fill = c("#D55E00", "#66CC99"), alpha = 0.7),
     labels = list(font = 2),
     quantities = TRUE)

all_proteins <- unique(c(vir1, u2b))

overlap_dt2 <- data.table(
  protein = all_proteins,
  U2B = all_proteins %in% u2b,
  VIR = all_proteins %in% vir1
)
# Get mappings
gene_names <- AnnotationDbi::select(org.At.tair.db,
                                    keys = overlap_dt2$protein,
                                    columns = c("SYMBOL"),
                                    keytype = "TAIR")

# Merge
overlap_dt2 <- merge(overlap_dt2, gene_names, by.x = "protein", by.y = "TAIR", all.x = TRUE)
overlap_dt2 <- overlap_dt2[, .(
  SYMBOL = paste(unique(SYMBOL), collapse = ",")
), by = .(protein, U2B, VIR)]
# Subset proteins in all three
overlap_dt2[U2B & VIR]
write_csv(overlap_dt2, file = "../../Proteomics/u2b_vir_overlap.csv")
```




#MTA in HAKAI
```{r}
mta_vs_mtahakai$colour <- ""
mta_vs_mtahakai$colour[mta_vs_mtahakai$protein %in% writer_complex] <- 1
mta_vs_mtahakai$colour[mta_vs_mtahakai$protein == "AT5G01160"] <- 2
mta_vs_mtahakai$name[mta_vs_mtahakai$protein == "AT5G01160"] <- "HAKAI"
mta_vs_mtahakai$name[mta_vs_mtahakai$protein == "AT3G05680"] <- "VIR"
mta_vs_mtahakai$name[mta_vs_mtahakai$protein == "AT5G53440"] <- "HIZ2"
mta_vs_mtahakai$name[mta_vs_mtahakai$protein == "AT4G10760"] <- "MTA"
mta_vs_mtahakai$name[mta_vs_mtahakai$protein == "AT4G09980"] <- "MTB"
mta_vs_mtahakai$name[mta_vs_mtahakai$protein == "AT3G54170"] <- "FIP37"
mta_vs_mtahakai$name[mta_vs_mtahakai$protein == "AT1G32360"] <- "HIZ1"

plot_mta_vs_mtahakai <- ggplot(data = mta_vs_mtahakai[mta_vs_mtahakai$MTA>=3,], aes(y=-log10(pval), x=median_logFC, label = name, fill = colour, size = median_av_exprs))+
    geom_point(shape = 21,colour = "white")+
    theme_classic(base_size=15)+
    geom_text_repel(aes(label=name, hjust=1.2), size = 6, max.overlaps = Inf)+
    xlab("median logFC")+
    #xlim(-4,12)+
    xlab('Log2 Fold Change')+
    ylab('-Log10 P value')+
    ggtitle("MTA vs MTA in HAKAI IP")+
    scale_fill_manual(labels = c("", "mRNA binding"),  values = c("#B5B5B5", "#7f47ca", "#717171"), name="")+
    theme(legend.position = c(.8, .2))+
    guides(size = "none", fill = "none")

plot_mta_vs_mtahakai
```


Ect plots
```{r}
plot_A <- ggplot(data = prot_res_ect[prot_res_ect$ECT2>=3,], aes(y=-log10(pval), x=median_logFC, label = name, fill=rna, size = median_av_exprs))+
    geom_point(shape = 21,colour = "white")+
    theme_classic(base_size=15)+
    geom_text_repel(aes(label=name, hjust=1.2), size = 6, max.overlaps = Inf)+
    xlab("median logFC")+
    xlim(-4,12)+
    xlab('Log2 Fold Change')+
    ylab('-Log10 P value')+
    ggtitle("RNA binding proteins")+
    scale_fill_manual(labels = c("", "mRNA binding"),  values = c("#B5B5B5", "#7f47ca", "#717171"), name="")+
    theme(legend.position = c(.8, .2))+
    guides(size = "none", fill = "none")

plot_A
```
Ect ranked plot
```{r}
# Add rank column
prot_res_ect[, logFC_rank := rank(-median_logFC)]

# Create the plot
p <- ggplot(prot_res_ect, aes(x = logFC_rank, y = median_logFC)) +
  geom_point(alpha = 0.6, color = "darkblue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") +
  geom_text_repel(
    data = prot_res_ect,
    aes(label = name),
    size = 3,
    max.overlaps = Inf,
    color = "firebrick"
  ) +
  labs(
    title = "ECT2",
    x = "Enrichment Rank",
    y = "Median log2FC"
  ) +
  theme_classic()

# Show plot
print(p)

# Optionally save it
ggsave(
  filename = "../../Proteomics/plots/ect_labelled.pdf",
  plot = p,
  width = 6, height = 4
)
```

